{\normalsize Seminar: Formal Specification} \\[1ex] Security of Multithreaded Programs by Compilation\cite{Barthe07}

### 2012-01-29 paper/paper.tex
added section about observational determinism

\subsection{Observational Determinism for Concurrent Program Security\cite{Zdancewic03}}
\subsection{Flexible Scheduler-Independent Security\cite{Mantel10}}
This approach developed by Mantel and Sudbrock has the goal to be on the one hand
scheduler independent and on the other not too restrictive (i.e. does not treat
obviously secure programs as insecure).

>>>>

\subsection{Observational Determinism for Concurrent Program Security\cite{Zdancewic03}}
Zdancewic and Myers approach has to goal to provide a secure concurrent language
that has a general and realistic support for concurrency and whose security can
be checked statically.

They introduced the language $\lambda_{SEC}^{PAR}$ which supports higher-order functions, an
unbounded number of threads, synchronization (via join patterns), message passing and shared memory. Regardless of those realistic
features $\lambda_{SEC}^{PAR}$ is not intended to server as a user-lever programming language because its
syntax and type system is too awkward, instead it is a model for studying information flow.

Low-security observational determinism is used to ensure noninterference. The idea is
that to make the naturally nondeterministic system, deterministic from the point of view
of the attacker.

This deterministic noninterference is defined as

\begin{align*}
(m \approx_\zeta m' \land m \Downarrow T \land m' \Downarrow T') \Rightarrow T \approx_\zeta T'
\end{align*}

Where $T$ and $T'$ are traces of the execution of a program. $m$ and $m'$ are initial configurations.
Let $T(L) = [M_0(L), M_1(L), \dots]$ be the list of values at location $L$ in the trace $T$. Then
$T \approx_\zeta T'$ holds if the values of $T(L)$ and $T'(L)$ are pairwise indistinguishable. Since
it is allowed that one sequence is a prefix of the other, there is no need to proof termination, but
external timing attacks are possible.

This notion of observational determinism is captured in a type system for $\lambda_{SEC}^{PAR}$, thus
the security of programs can be checked statically.

\subsection{Flexible Scheduler-Independent Security\cite{Mantel10}}
This approach developed by Mantel and Sudbrock has the goal to be on the one hand
scheduler independent and on the other not too restrictive (i.e. does not treat
obviously secure programs as insecure).
